Oxidation Protocol

### 2019-11-01 R/process.Rmd
Save before switching to first-order decay

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,  fig.align="center")
library(tidyverse)
library(knitr)
library(kableExtra)
library(broom)
library(modelr)
library(lubridate)
```

@@

```{r sc_data, fig.height=3}
example_sc <- data.frame(
  flux_date = c('2019-05-27', '2019-05-27', '2019-05-27', '2019-06-07', '2019-06-07', '2019-06-07', '2019-06-14', '2019-06-14', '2019-06-14'),
  standard_ppm = c(999, 999, 999, 999, 999, 999, 999, 999, 999),
  injection_ml = c(0.1, 0.5, 1, 0.1, 0.5, 1, 0.1, 0.5, 1),
  ch4_area = c(556.03, 3046.03, 5995.73, 493.256, 2720.52, 5259.6, 569.3, 3255.82, 6129.72))
example_sc <- 
  example_sc %>% 
  mutate(injection_ppm = standard_ppm * injection_ml) %>% 
  select(flux_date:injection_ml, injection_ppm, ch4_area)
example_sc %>% 
  kable() %>% 
  kable_styling()
ggplot(example_sc, aes(ch4_area, injection_ppm)) +
  facet_wrap(~ flux_date) + 
  geom_point() +
  geom_line() +
  coord_fixed(6)

sc_model <- function(df) {lm(injection_ppm ~ ch4_area, data = df)}
standard_curves <- 
  example_sc %>% 
  group_by(flux_date) %>% 
  nest() %>% 
  mutate(model = map(data, sc_model),
         glance = map(model, glance),
         tidy = map(model, tidy)) 
sc_r2 <- 
  standard_curves %>% 
  unnest(glance, .drop = TRUE) %>% 
  arrange(r.squared) %>% 
  select(flux_date, r.squared)
sc_equation <- 
  standard_curves %>% 
  unnest(tidy, .drop = TRUE) %>%
  select(flux_date, term, estimate) %>% 
  spread(term, estimate) %>% 
  rename(intercept = `(Intercept)`, slope = ch4_area)
sc_ch4 <- left_join(sc_r2, sc_equation, by = "flux_date")
sc_ch4 %>% 
  kable() %>% 
  kable_styling()
```

@@

time_data <- data.table::fread('../Data/eg_time_data.csv')

time_data[, c('t0', 't1', 't2', 't3', 't4')] <- 
  lapply(time_data[, c('t0', 't1', 't2', 't3', 't4')],
         ymd_hms)

@@

time_data[, c('t1', 't2', 't3', 't4')] <- 
  lapply(time_data[, c('t1', 't2', 't3', 't4')],
         function(x) as.days(time_data$t0, x))

time_data <- 
 time_data %>% 
  select(-t0) %>% 
  gather(t, days, t1:t4)

@@

conc_data <- data.table::fread('../Data/eg_conc_data.csv')

flux_data <- 
  conc_data %>% 
  gather(t, ch4_area, t1:t4) %>% 
  left_join(sc_ch4, by = 'flux_date') %>% 
  mutate(ch4_ppm = ch4_area * slope + intercept) %>% 
  select(flux_date, jar, t, ch4_ppm) %>% 
  left_join(time_data, by = c('flux_date', 'jar', 't'))

@@

```{r plot_fluxes, fig.height=20}
flux_data <- 
  flux_data %>% 
  mutate(flux_date = ymd(flux_date), 
         jar = factor(jar),
         t = factor(t)) %>% 
  filter(ch4_ppm < 3000)

flux_data %>% 
  ggplot(aes(days, ch4_ppm)) + 
  facet_grid(jar ~ flux_date, scales = "free") +
  geom_point() + 
  geom_line()
```

### Calculate Flux

```{r calc_flux}

@@

```{r conversion}

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,  fig.align="center")
library(tidyverse)
library(knitr)
library(kableExtra)
library(broom)
library(modelr)
library(lubridate)
library(morris)
```

@@

```{r sc_data, fig.height=3}

sc <- read.csv('../Data/standard_curve.csv')

ggplot(sc, aes(area, injection_ppm)) +
  facet_wrap(~ flux_date + molecule, scales = "free") + 
  geom_point() +
  stat_smooth(method = 'lm', se = FALSE)

sc_model <- function(df) {lm(injection_ppm ~ area, data = df)}
standard_curves <- 
  sc %>% 
  group_by(flux_date, molecule) %>% 
  nest() %>% 
  mutate(model = map(data, sc_model),
         glance = map(model, glance),
         tidy = map(model, tidy)) 
sc_r2 <- 
  standard_curves %>% 
  unnest(glance, .drop = TRUE) %>% 
  arrange(r.squared) %>% 
  select(flux_date, molecule, r.squared)
sc_equation <- 
  standard_curves %>% 
  unnest(tidy, .drop = TRUE) %>%
  select(flux_date, molecule, term, estimate) %>% 
  spread(term, estimate) %>% 
  rename(intercept = `(Intercept)`, slope = area)
sc_ch4 <- left_join(sc_r2, sc_equation, by = c("flux_date", "molecule"))
sc_ch4 %>% 
  kable() %>% 
  kable_styling()

@@

time_data <- data.table::fread('../Data/time_data.csv')

time_data[, c('t0', 't1', 't2', 't3', 't4', 't5')] <- 
  lapply(time_data[, c('t0', 't1', 't2', 't3', 't4', 't5')],
         ymd_hms)

@@

time_data[, c('t1', 't2', 't3', 't4', 't5')] <- 
  lapply(time_data[, c('t1', 't2', 't3', 't4', 't5')],
         function(x) as.days(time_data$t0, x))

time_data <- 
 time_data %>% 
  select(-t0) %>% 
  gather(t, days, t1:t5)

@@

conc_data <- data.table::fread('../Data/conc_data.csv')

flux_data <- 
  conc_data %>% 
  gather(t, area, t1:t5) %>% 
  left_join(sc_ch4, by = c('flux_date', 'molecule')) %>% 
  mutate(ppm = area * slope + intercept) %>% 
  select(flux_date, jar, molecule, t, ppm)  %>% 
  left_join(time_data, by = c('flux_date', 'jar', 't'))

@@

```{r plot_fluxes}
flux_data <- 
  flux_data %>% 
  mutate(flux_date = factor(ymd(flux_date)), 
         jar = factor(jar),
         molecule = factor(molecule),
         t = factor(t))

flux_data <- 
  flux_data %>% 
  mutate(rep = factor(substr(jar, 2, 3)),
         treat = factor(substr(jar, 1, 1)))

# P1: CO2

```{r, fig.height = 30, warning = FALSE}

flux_data %>% 
  filter(flux_date == '2019-09-24') %>% 
  filter(molecule == 'co2') %>% 
  ggplot(aes(days, ppm)) + 
  facet_wrap(c('treat', 'rep'), scales = "free", ncol = 2) +
  geom_point() + 
  stat_smooth(method = "nls", formula = y ~ SSasymp(x, Asym, R0, lrc), se = FALSE) +
  expand_limits(y = 0)

# P1: CH4

```{r, fig.height = 30, warning = FALSE}

flux_data %>% 
  filter(flux_date == '2019-09-24') %>% 
  filter(molecule == 'ch4') %>% 
  ggplot(aes(days, ppm)) + 
  facet_wrap(c('treat', 'rep'), scales = "free", ncol = 2) +
  geom_point() + 
  stat_smooth(method = "nls", formula = y ~ SSasymp(x, Asym, R0, lrc), se = FALSE) +
  expand_limits(y = 0)

# P2: CO2

```{r, fig.height = 30, warning = FALSE}

flux_data %>% 
  filter(flux_date == '2019-10-23') %>% 
  filter(molecule == 'co2') %>% 
  ggplot(aes(days, ppm)) + 
  facet_wrap(c('treat', 'rep'), scales = "free", ncol = 2) +
  geom_point() + 
  stat_smooth(method = "nls", formula = y ~ SSasymp(x, Asym, R0, lrc), se = FALSE) +
  expand_limits(y = 0)

# P2: CH4

```{r, fig.height = 30, warning = FALSE}

flux_data %>% 
  filter(flux_date == '2019-10-23') %>% 
  filter(molecule == 'ch4') %>% 
  ggplot(aes(days, ppm)) + 
  facet_wrap(c('treat', 'rep'), scales = "free", ncol = 2) +
  geom_point() + 
  stat_smooth(method = "nls", formula = y ~ SSasymp(x, Asym, R0, lrc), se = FALSE)

# P2: CH4 baseline 0

```{r, fig.height = 30, warning = FALSE}

flux_data %>% 
  filter(flux_date == '2019-10-23') %>% 
  filter(molecule == 'ch4') %>% 
  ggplot(aes(days, ppm)) + 
  facet_wrap(c('treat', 'rep'), scales = "free", ncol = 2) +
  geom_point() + 
  stat_smooth(method = "nls", formula = y ~ SSasymp(x, Asym, R0, lrc), se = FALSE) +
  expand_limits(y = 0)

```{r flux_model}
lin_model <- function(dt) {
  lm(ppm ~ days, data = dt)
}
exp_model <- function(dt) {
  t <- try(nls(ppm ~ SSasymp(days, Asym, R0, lrc), data = dt))
  if(inherits(t, "try-error")) return('foo') else t
}

nested <- 
flux_data %>% 
  group_by(flux_date, jar, molecule, treat, rep) %>% 
  nest() %>% 
  mutate(model = map(data, exp_model))

lin_nested <- 
flux_data %>% 
  group_by(flux_date, jar, molecule, treat, rep) %>% 
  nest() %>% 
  mutate(model = map(data, lin_model))

fluxes <-
  nested %>% 
  mutate(class = map(model, class)) %>% 
  unnest(class) %>% 
  filter(class == 'nls') %>% 
  mutate(tidy = map(model, tidy)) %>% 
  unnest(tidy) %>% 
  filter(term == 'lrc') %>% 
  filter(molecule == 'ch4') %>% 
  mutate(lambda = exp(estimate)) %>% 
  group_by(flux_date) %>% 
  arrange(lambda, .by_group = TRUE) %>% 
  select(-data, -model, -class, -term)

fluxes %>% 
  kable() %>% 
  kable_styling()

lin_fluxes <- 
lin_nested %>% 
  mutate(tidy = map(model, tidy)) %>% 
  unnest(tidy) %>% 
  filter(term == 'days')  %>% 
  group_by(flux_date) %>% 
  arrange(estimate, p.value, .by_group=TRUE) %>% 
  select(-data, -model)

lin_fluxes %>% 
  kable() %>% 
  kable_styling()
```

```{r test_flux, eval = FALSE}

summary(aov(lambda ~ treat + flux_date, data = fluxes))
flux_data
nls(ppm ~ SSasymp(days, Asym, R0, lrc), data = flux_data)
nested
fluxes %>% 
  ggplot(aes(x = lambda, fill = treat)) +
  facet_wrap(~ flux_date) +
  geom_histogram(binwidth = 1, alpha = 0.5, position = 'identity')

lin_fluxes %>% 
  filter(molecule == 'ch4') %>% 
  ggplot(aes(x = estimate, fill = treat)) +
  facet_wrap(~ flux_date) +
  geom_histogram(alpha = 0.5, position = 'identity')

```{r plot_resids, fig.height=20, eval=FALSE}

by_date_jar %>% 
  mutate(model = map(data, flux_model)) %>% 
  mutate(resids = map2(data, model, add_residuals)) %>% 
  unnest(resids) %>% 
  ggplot(aes(days, abs(resid), group = interaction(flux_date, jar))) +
  facet_grid(jar ~ flux_date, scales = "free") +
  geom_point() + 
  geom_line()

by_date_jar %>% 
  unnest(data) %>% 
  summarize(length(t)) %>% 
  print(n = 100)

@@

```{r conversion, eval = FALSE}

all_data <- read.csv('../Data/all_data.csv')
head(all_data)
all_data %>% 
  gather(t, area, t1:t5) %>% 
  gather(t2, time, `t1.1`:`t5.1`) %>% 
  head()
tail(all_data)
flux_data %>% 
  arrange(flux_date, molecule, jar, t) %>% 
write.csv(., '../Data/all_data.csv')
```
